---
title: "portfolio 3"
author: "Sofie Mosegaard"
date: "2022-11-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
pacman::p_load(msm, tidyverse, brms, grid, gridExtra, readxl, metafor, dplyr, magrittr, reshape, tidymodels)

```

The Machine Learning assignment has 3 main parts: First we create a skeptical and an informed simulation, based on the meta-analysis. Second we build and test our machine learning pipeline on the simulated data. Second we apply the pipeline to the empirical data.

The report for the exam, thus, consists of the answer to all the following prompts:
- Describe your machine learning pipeline. Produce a diagram of it to guide the reader (e.g. see Rybner et al 2022 Vocal markers of autism: Assessing the generalization of ML models), and describe the different parts: data budgeting, data preprocessing, model choice and training, assessment of performance.
- Briefly justify and describe your use of simulated data, and results from the pipeline on them.
- Describe results from applying the ML pipeline to the empirical data and what can we learn from them.

Remember: plots are very very important to communicate your process and results.

### Part I - Simulating data

Use the meta-analysis reported in Parola et al (2020), create a simulated dataset with 100 matched pairs of schizophrenia and controls, each participant producing 10 repeated measures (10 trials with their speech recorded). for each of these "recordings" (data points) produce 10 acoustic measures: 6 from the meta-analysis, 4 with just random noise. Do the same for a baseline dataset including only 10 noise variables. Tip: see the slides for the code. 

#Simulating data (df 1 = with 6 from the meta-analysis, 4 with just random noise)
```{r}

# Define population size
n <- 100
trials <- 10

# Define effect sizes
#see table 2 from the article for the numbers
#the last 4 numbers are 0, as dont want any noice in the underlying mechanisms 
InformedEffectMean <- c(0.25, -0.55, 0.74, -1.26, 0.05, 1.89, 0, 0, 0, 0)

# Define individual variability from population and across trails and measurement error
IndividualSD <-  1
TrialSD <- 0.5
Error <- 0.2

# Conditions
Schizophrenia <- rnorm(1, rnorm(1, 0.21, 0.5))
Control <- rnorm(1, rnorm(1, -0.21, 0.5), 0.2)

# For each pair of participant, we need to identity the true effect size
for(i in seq(10)) {
  temp_informed <- tibble(
    ID = seq(n),
    TrueEffect = rnorm(n, InformedEffectMean[i], IndividualSD),
    Variable = paste0("v", i))
  if(i == 1) {
    d_informed_true <- temp_informed
  } else {
    d_informed_true <- rbind(d_informed_true, temp_informed)
  }
}

#(temp = temperately)

# Create tibble with one row per trial
d_trial <- tibble(expand_grid(ID = seq(n), Trial = seq(trials), Group = c("Schizophrenia", "Control")))

d_informed <- merge(d_informed_true, d_trial)

for(i in seq(nrow(d_informed))){
  d_informed$measurement[i] <- ifelse(d_informed$Group[i]=="Schizophrenia",
                                      rnorm(1, rnorm(1, d_informed$TrueEffect[i]/2, TrialSD), Error),
                                      rnorm(1, rnorm(1, -d_informed$TrueEffect[i]/2, TrialSD), Error))
}

#from row values to columns - look at the diff between "d_informed_trail" and "..._wide"
d_informed_wide <- d_informed %>% 
  mutate(TrueEffect= NULL) %>% 
  pivot_wider(names_from = Variable,
              values_from = measurement)

d_informed_wide <- d_informed_wide %>% 
  mutate(ID = as.factor(ID)) %>% 
  mutate(Trial = as.factor(Trial))

```

#second simulation (df 2 = baseline dataset including only 10 noise variables)
```{r}
#... and now we do the same for the second data set (baseline dataset including only 10 noise variables)

skeptic_EffectMean <- rep(0,10)

for(i in seq(10)) {
  temp_skeptic <- tibble(
    ID = seq(n),
    TrueEffect = rnorm(n, skeptic_EffectMean[i], IndividualSD),
    Variable = paste0("v", i))
  if(i == 1) {
    d_informed_true <- temp_informed
    d_skeptic_true <- temp_skeptic
  } else {
    d_skeptic_true <- rbind(d_skeptic_true, temp_skeptic)
  }
}

d_skeptic <- merge(d_skeptic_true, d_trial)

for(i in seq(nrow(d_skeptic))){
  d_skeptic$measurement[i] <- ifelse(d_skeptic$Group[i]=="Schizophrenia",
                                      rnorm(1, rnorm(1, d_skeptic$TrueEffect[i]/2, TrialSD), Error),
                                      rnorm(1, rnorm(1, -d_skeptic$TrueEffect[i]/2, TrialSD), Error))
}
d_skeptic_wide <- d_skeptic %>%
  mutate(TrueEffect= NULL) %>% 
  pivot_wider(names_from = Variable,
              values_from = measurement)

d_skeptic_wide <- d_skeptic_wide %>%
    mutate(ID = as.factor(ID)) %>%
    mutate(Trial = as.factor(Trial))
```


### Part II - ML pipeline on simulated data

On the two simulated datasets (separately) build a machine learning pipeline: i) create a data budget (e.g. balanced training and test sets); ii) pre-process the data (e.g. scaling the features); iii) fit and assess a classification algorithm on the training data (e.g. Bayesian multilevel logistic regression); iv) assess performance on the test set; v) discuss whether performance is as expected and feature importance is as expected.

Bonus question: replace the bayesian multilevel regression with a different algorithm, e.g. SVM or random forest (but really, anything you'd like to try).

# i) create a data budget (e.g. data budgeting)
```{r}

TestID <- sample(seq(n), 20)

train_informed <- d_informed_wide %>% 
  subset(!(ID %in% TestID))
test_informed <- d_informed_wide %>% 
  subset(ID %in% TestID)

train_skeptic <- d_skeptic_wide %>% 
  subset(!(ID %in% TestID))
test_skeptic <- d_skeptic_wide %>% 
  subset((ID %in% TestID))

```

# ii) pre-process the data (e.g. scaling the features)
```{r}

#informed

scaled_informed <- train_informed %>% 
  recipe(Group ~ .) %>% #define the outcome
  step_scale(all_numeric()) %>% #scales all the numeric values
  step_center(all_numeric()) %>% #centers all the numeric values
  prep(training = train_informed, retain = TRUE)

  
#creating new scaled dataframes
train_informed_scaled <- juice(scaled_informed) 
test_informed_scaled <- bake(scaled_informed, new_data = test_informed)


#skeptic
scaled_skeptic <- train_skeptic %>% 
  recipe(Group ~ .) %>% #define the outcome
  step_scale(all_numeric()) %>% #scales all the numeric values
  step_center(all_numeric()) %>% #centers all the numeric values
  prep(training = train_skeptic, retain = TRUE)

#creating new scaled dataframes
train_skeptic_scaled <- juice(scaled_skeptic)
test_skeptic_scaled <- bake(scaled_skeptic, new_data = test_skeptic)
```

#iii) fit and assess a classification algorithm on the training data (e.g. Bayesian multilevel logistic regression)
```{r}
#create formula for informed
f1 <- bf(Group ~ 1 + v1 + v2 + v3 + v4 + v5 +v6 +v7 +v8 +v9 +v10)

#create formula for skeptic
s_f1 <- bf(Group ~ 1 + v1 + v2 + v3 + v4 + v5 +v6 +v7 +v8 +v9 +v10)

```
#get prios
```{r}
#informed
get_prior(f1, 
          train_informed_scaled, 
          family = bernoulli)

#skeptic
get_prior(s_f1, 
    train_skeptic_scaled,
    family = bernoulli)
```

#setting priors
```{r}
#informed
f1_prior <- c(
  prior(normal(0, 1), class = Intercept),
  prior(normal(0, 0.3), class = b) 
  )

#skeptic
s_f1_prior <- c(
prior(normal(0, 1), class = Intercept),
prior(normal(0, 0.3), class = b) 
)
```

### Make model
```{r}
#informed
fitted_f1_prior <- brm(
    f1,
    train_informed_scaled,
    family = bernoulli,
    prior = f1_prior,
    sample_prior = T,
    iter = 4000, 
    warmup = 2000,
    cores = 4,
    backend ="cmdstanr",
    threads = threading(2),
    refresh=0,
    chains = 4,
    control = list(
      adapt_delta = 0.999,
      max_treedepth = 20))

#skeptic
fitted_s_f1_prior <- 
  brm(
    s_f1,
    train_skeptic_scaled,
    family = bernoulli,
    prior = s_f1_prior,
    sample_prior = T,
    iter = 4000, 
    warmup = 2000,
    cores = 4,
    backend ="cmdstanr",
    threads = threading(2),
    refresh=0,
    chains = 4,
    control = list(
      adapt_delta = 0.999,
      max_treedepth = 20))

```

#look at pp-check
```{r}
#informed
pp_check(fitted_f1_prior, ndraws = 100)

#skeptic
pp_check(fitted_s_f1_prior, ndraws = 100)
```

# Parameter recovery
```{r}
#informed
print(fitted_f1_prior)

#skeptic
print(fitted_s_f1_prior)
```

#Prior posterier update checks
```{r}

#informed
variables(fitted_f1_prior)
Posterior_f1 <- as_draws_df(fitted_f1_prior)

ggplot(Posterior_f1) +
  geom_density(aes(prior_Intercept), fill="chartreuse2", color="black",alpha=0.6) +
  geom_density(aes(b_Intercept), fill="deeppink", color="black",alpha=0.6) + 
  xlab('Intercept') +
  theme_classic()+
  ggtitle("intercept for study effect size")+
      theme(plot.title = element_text(size = 10, face = "bold"))

#skeptic
variables(fitted_s_f1_prior)
Posterior_s_f1 <- as_draws_df(fitted_s_f1_prior)

ggplot(Posterior_s_f1) +
  geom_density(aes(prior_Intercept), fill="chartreuse2", color="black",alpha=0.6) +
  geom_density(aes(b_Intercept), fill="deeppink", color="black",alpha=0.6) + 
  xlab('Intercept') +
  theme_classic()+
  ggtitle("intercept for study effect size")+
      theme(plot.title = element_text(size = 10, face = "bold"))


#green = prior
#pink = simulated data 
```
  
iv) assess performance on the test set; 
```{r}
PerformanceProb <- tibble(expand_grid(
    Sample = seq(4000),
    Setup= c("informed", "skeptic"),
    Type= c("training", "test"))
    )

# informed
test_informed_scaled$PredictionsPerc0 <- predict(fitted_f1_prior, newdata = test_informed_scaled, allow_new_levels = T)[,1]
test_informed_scaled$Predictions0[test_informed_scaled$PredictionsPerc0 > 0.5] <- "Schizophrenia" 
test_informed_scaled$Predictions0[test_informed_scaled$PredictionsPerc0 <= 0.5] <- "Control" 

train_informed_scaled$PredictionsPerc0 <- predict(fitted_f1_prior)[,1]
train_informed_scaled$Predictions0[train_informed_scaled$PredictionsPerc0 > 0.5] <- "Schizophrenia" 
train_informed_scaled$Predictions0[train_informed_scaled$PredictionsPerc0 <= 0.5] <- "Control" 

# skeptic
test_skeptic_scaled$PredictionsPerc1 <- predict(fitted_s_f1_prior, newdata = test_skeptic_scaled, allow_new_levels = T)[,1]
test_skeptic_scaled$Predictions1[test_skeptic_scaled$PredictionsPerc1 > 0.5] <- "Schizophrenia" 
test_skeptic_scaled$Predictions1[test_skeptic_scaled$PredictionsPerc1 <= 0.5] <- "Control" 

train_skeptic_scaled$PredictionsPerc1 <- predict(fitted_s_f1_prior)[,1]
train_skeptic_scaled$Predictions1[train_skeptic_scaled$PredictionsPerc1 > 0.5] <- "Schizophrenia" 
train_skeptic_scaled$Predictions1[train_skeptic_scaled$PredictionsPerc1 <= 0.5] <- "Control" 


# informed 
train0 <- inv_logit_scaled(posterior_linpred(fitted_f1_prior, 
    summary = F))
    
test0 <- inv_logit_scaled(posterior_linpred(fitted_f1_prior, 
    summary = F, 
    newdata = test_informed_scaled, 
    allow_new_levels = T ))
    
# skeptic
train1 <- inv_logit_scaled(posterior_linpred(fitted_s_f1_prior, 
    summary = F))
    
test1 <- inv_logit_scaled(posterior_linpred(fitted_s_f1_prior, 
    summary = F, 
    newdata = test_skeptic_scaled, 
    allow_new_levels = T ))

# informed
test_informed_scaled <- test_informed_scaled %>% 
  mutate(Group = as.factor(Group),
         Predictions0 = as.factor((Predictions0)))

train_informed_scaled <- train_informed_scaled %>% 
  mutate(Group = as.factor(Group),
         Predictions0 = as.factor((Predictions0)))

# skeptic
test_skeptic_scaled <- test_skeptic_scaled %>% 
  mutate(Group = as.factor(Group),
         Predictions1 = as.factor(Predictions1))

train_skeptic_scaled <- train_skeptic_scaled %>% 
  mutate(Group = as.factor(Group),
         Predictions1 = as.factor(Predictions1))



# obs the following for-loop does not work:

for (i in seq(4000)){

    train_informed_scaled$Predictions0 <- as.factor(ifelse(train0[i,] > 0.5, "Schizophrenia", "Control"))
    
    test_informed_scaled$Predictions0 <- as.factor(ifelse(test0[i,] > 0.5, "Schizophrenia", "Control"))

    PerformanceProb$Accuracy[PerformanceProb$Sample==i & PerformanceProb$Setup == "informed" & PerformanceProb$Type =="training"] <- 
    accuracy(train_informed_scaled, truth = Group, estimate = Predictions0)[, ".estimate"]
    
    PerformanceProb$Accuracy[PerformanceProb$Sample==i & PerformanceProb$Setup == "informed" & PerformanceProb$Type =="test"] <- 
    accuracy(test_informed_scaled, truth = Group, estimate = Predictions0)[, ".estimate"]
    
    train_skeptic_scaled$Predictions1 <- as.factor(ifelse(train1[i,] > 0.5, "Schizophrenia", "Control"))
    test_skeptic_scaled$Predictions1 <- as.factor(ifelse(test1[i,] > 0.5, "Schizophrenia", "Control"))
  
    PerformanceProb$Accuracy[PerformanceProb$Sample == i & PerformanceProb$Setup == "skeptic" & PerformanceProb$Type =="training"] <- 
    accuracy(train_skeptic_scaled, truth = Group, estimate = Predictions1)[, ".estimate"]
    
    
    PerformanceProb$Accuracy[PerformanceProb$Sample == i & PerformanceProb$Setup == "skeptic" & PerformanceProb$Type =="test"] <- 
    accuracy(test_skeptic_scaled, truth = Group, estimate = Predictions1)[, ".estimate"]
}

```

```{r}
# Assessing (average performance)
conf_mat(
  test_informed_scaled,
  truth = Group,
  estimate = Predictions0,
  dnn = c("Predictions", "Truth")
)

metrics(test_informed_scaled,
        truth = Group, estimate = Predictions0) %>% 
  knitr::kable()
```

```{r}
#Plotting the accuracy

ggplot(PerformanceProb) +
  geom_point(aes(x = Setup, y = as.numeric(Accuracy), colour = Type)) + geom_abline(intercept = 0.5, slope = 0, col=c("Purple"), linetype = c("dashed")) +
  theme_minimal() +
  ylab("Accuracy") +
  xlab("Type") + 
  theme_minimal() +
  ggtitle("Accuracy between informed and sceptic")

```


v) discuss whether performance is as expected and feature importance is as expected.


### Part III - Applying the ML pipeline to empirical data

Download the empirical dataset from brightspace and apply your ML pipeline to the new data, adjusting where needed. Warning: in the simulated dataset we only had 10 features, now you have many more! Such is the life of the ML practitioner. Consider the impact a higher number of features will have on your ML inference, and decide whether you need to cut down the number of features before running the pipeline (or alternatively expand the pipeline to add feature selection).

#load data
```{r}
data <- read_csv("Ass3_empiricalData1.csv")
```

#Making the formula
```{r}
pacman::p_load(DALEX, DALEXtra, kernlab, randomForest, xgboost, knitr, dotwhisker)

d_inf <- train_informed_scaled %>%
        mutate(ID = NULL, Trial = NULL, Preds = NULL, Predictions = NULL, v1_s = NULL)
        
LogisticRegression_inf <- logistic_reg() %>%
        set_mode("classification") %>%
        set_engine("glm") %>%
        fit(Group ~., data=d_inf)

explainer_lm <- explain_tidymodels(
        LogisticRegression_inf,
        data = train_informed_scaled,
        y = as.numeric(train_informed_scaled$Group) -1,
        label = "logReg",
        verbose = FALSE)

explainer_lm %>% 
        model_parts() %>%
        plot(show_boxplots = FALSE) +
        ggtitle("Feature Importance", "")

explainer_rf <- explain_tidymodels(
        RandomForest_inf,
        data = train_informed_scaled,
        y = as.numeric(train_informed_scaled$Group) -1,
        label = "random forest",
        verbose = FALSE)

explainer_rf %>% 
        model_parts() %>%
        plot(show_boxplots = FALSE) +
        ggtitle("Feature Importance", "")
```

#Get prior
```{r}

```

#Set priors
```{r}

```

#Building model
```{r}
```


```{r}
#Prior predictive checks
```


```{r}
#Fit the model
```

```{r}
#Posterior predictive checks
```


```{r}
#Prior posterior update checks 
```




